---
title: "Using `map` and `walk` from the `purrr` package"
author: Bolívar Aponte Rolón, Ph.D.
date: "2024-11-08"
format: 
  html:
    toc: true
    toc-location: left
    toc-depth: 2
    number-sections: true
    number-depth: 1
    theme: lumen
    highlight-style: github
    code-overflow: wrap
    code-fold: false
    code-copy: true
    code-link: true
    code-tools: false
    code-block-border-left: "#0C3823"
    code-block-bg: "#eeeeee"
    fig-cap-location: margin
    linestretch: 1.25
    fontsize: "large"
    embed-resources: true
editor_options: 
  
  chunk_output_type: console
execute:
    echo: true
    eval: false
---

## Setup

```{r}
#| label: setup


library(tidyverse)
library(gapminder)
```

## Exercise 1

-   Use `map()` and `lm_ggplot` to produce a *list* of scatter plots for the countries below

-   Save it to an object called `plots`

```{r}
countries <- c("Afghanistan",
               "United States",
               "United Kingdom",
               "China",
               "India")
```

```{r}
# Fill in the missing parts to save the plots

____ <- map(countries,
             \(.x) ______(
               data = _____,
               ______,
               x = year,
               y = ______
             ))
```

## Exercise 2

Use `walk2()` to save your *ggplots*:

Setting up where to save them

```{r}
# make the folder to put them it (if exists, {fs} does nothing)
fs::dir_create("plots")

# make the file names
plot_paths <- str_glue("plots/{print(countries)}.png")
```

::: callout-tip
Usage of `::` helps to access the exact function from that specific package. This avoids conflicts with other packages.
:::

::: callout-note
Check out the [*fs*](https://fs.r-lib.org/index.html) and [*glue*](https://glue.tidyverse.org/) packages for more information on the functions we used above.
:::

You can check the documentation on the [`purrr`](https://purrr.tidyverse.org/) package for further details

```{r}
#| eval: false

# Fill in the missing parts to save the plots
walk2(
  ______,
  ______,
  \(path, plot) ______
)
```

Let's check if they are saved:

```{r}
#| eval: true
fs::dir_tree("plots")
```

## Resources and Further Reading

-   [R for Data Science](https://r4ds.had.co.nz/)
-   [Functional Programming in R](https://dcl-prog.stanford.edu/)
-   [Advanced R](https://adv-r.hadley.nz/index.html)
-   [Gapminder vignette](https://cran.r-project.org/web/packages/gapminder/vignettes/gapminder.html)
-   [Code Smells and Feels](https://github.com/jennybc/code-smells-and-feels)
-   [Efficiency and Consistency: Automate Subset Graphics with ggplot2 and purrr](https://www.cedricscherer.com/2023/07/05/efficiency-and-consistency-automate-subset-graphics-with-ggplot2-and-purrr/)
-   [`purrr::walk()` this way](https://www.tidyverse.org/blog/2023/05/purrr-walk-this-way/)
-   [Functional Programming in R](https://dcl-prog.stanford.edu/)
